---
title: "R project"
author: "Lee Shan Yan, Lee Zi Hooi"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Implementation of code

## Installing necessary libraries

```{r}
library(treemap)
library(dplyr)
library(ggplot2)
library(GGally)
library(corrplot)
library(rstatix)
library(broom)
library(kableExtra)
library(patternplot)
library(gganimate)
library(gapminder)
library(gridExtra) 
library(scales)
library(reshape2)
library(tidyr)
library(maps)

```

## Data processing

### Understand the data

```{r}
# Read data
billionaires <- read.csv("Billionaires Statistics Dataset.csv", header = TRUE)

str(billionaires)
```

## Dataset selection and transformation

```{r}
# Aggregate data by country
country_billionaire <- billionaires %>%
  group_by(country) %>%
  summarise(
    number_of_billionaires = n(),
    avg_final_worth = mean(finalWorth, na.rm = TRUE),
    total_final_worth = sum(finalWorth, na.rm = TRUE),
    most_common_industry = names(sort(table(industries), decreasing = TRUE)[1]),
    industry_diversity = n_distinct(industries),
    concatenated_industries = paste(industries, collapse = ", "),
    proportion_self_made = mean(selfMade, na.rm = TRUE), # Proportion of self-made billionaires
    proportion_male = mean(gender == "M", na.rm = TRUE), # Proportion of male billionaires
    cpi_country = first(cpi_country),
    cpi_change_country = first(cpi_change_country),
    gdp_country = first(gdp_country),
    gross_tertiary_education_enrollment = first(gross_tertiary_education_enrollment),
    gross_primary_education_enrollment_country = first(gross_primary_education_enrollment_country),
    life_expectancy_country = first(life_expectancy_country),
    tax_revenue_country = first(tax_revenue_country_country),
    total_tax_rate_country = first(total_tax_rate_country),
    population_country = first(population_country),
    latitude_country = first(latitude_country),
    longitude_country = first(longitude_country),
    .groups = 'drop'
  )

# Check structure
str(country_billionaire)
```

### Data type conversion

```{r}
# Remove dollar signs, commas and any whitespace for gdp_country
country_billionaire$gdp_country <- gsub("\\$|,|\\s+", "", country_billionaire$gdp_country)

# Convert cleaned gdp_country (char) to numeric
country_billionaire$gdp_country <- as.numeric(country_billionaire$gdp_country)
```

### Handling missing values: Checking

```{r}
# Check for missing values in all columns
print(colSums(is.na(country_billionaire)))
```

```{r}
unique(country_billionaire$country)
```

### Handling missing values

```{r}
# Remove rows where the country column is empty or missing
country_billionaire <- country_billionaire %>%
  filter(country != "" & !is.na(country))

unique(country_billionaire$country)
```

```{r}
# Function to calculate the mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Handle missing values for both numeric and categorical columns
impute_missing <- function(df) {
  for (col in names(df)) {
    if (is.numeric(df[[col]]) || is.integer(df[[col]])) {
      # For numeric columns, replace NA with the mean
      df[[col]][is.na(df[[col]])] <- mean(df[[col]], na.rm = TRUE)
    } else if (is.character(df[[col]])) {
      # For categorical columns, replace NA with the mode
      df[[col]][is.na(df[[col]])] <- Mode(df[[col]])
    }
  }
  return(df)
}
```

```{r}
# Apply the function to the billionaires dataset
new_billionaires_clean <- impute_missing(country_billionaire)

print(colSums(is.na(new_billionaires_clean)))
```

### Handling duplicates: Checking

```{r}
# Check for duplicates
duplicated(new_billionaires_clean)  # Identify duplicate rows
sum(duplicated(new_billionaires_clean))  
```

### View final cleaned dataset

```{r}
str(new_billionaires_clean)
```

```{r}
summary(new_billionaires_clean)
```

```{r}
head(new_billionaires_clean)
```

```{r}
View(new_billionaires_clean)
```

## Number of billionaires analysis

## Visualising countries-billionaires information

```{r}
# Load world map data
world_map <- map_data("world")

# Standardize country names in your dataset to match world_map
# If "USA" is how the United States is represented, map it accordingly
new_billionaires_clean$country <- dplyr::recode(
  new_billionaires_clean$country,
  "United States" = "USA"
)

# Check data types and convert to character if necessary
world_map$region <- as.character(world_map$region)
new_billionaires_clean$country <- as.character(new_billionaires_clean$country)

# Merge the datasets
geo_billionaire <- world_map %>%
  left_join(new_billionaires_clean, by = c("region" = "country"))

# Plot the map
ggplot() +
  geom_polygon(data = geo_billionaire, 
               aes(x = long, y = lat, group = group, fill = number_of_billionaires), 
               color = "grey80") +
  scale_fill_gradient(low = "pink", high = "purple", na.value = "white", name = "Number of Billionaires") +
  labs(title = "Global Distribution of Billionaires by Country",
       x = "Longitude", y = "Latitude") +
  coord_fixed(1.3) +  # Maintain aspect ratio
  theme_minimal() +
  theme(legend.position = "right")

```

```{r}
# Enhanced bar plot with shadow effect
library(ggplot2)
# 1. Top 10 Countries by Number of Billionaires
top_countries_billionaires <- new_billionaires_clean %>%
  arrange(desc(number_of_billionaires)) %>%
  slice(1:10)

plot_top_countries_billionaires <- ggplot(top_countries_billionaires) +
  # Shadow layer (slightly offset and darker)
  geom_bar(
    aes(x = reorder(country, -number_of_billionaires), y = number_of_billionaires),
    stat = "identity",
    width = 0.7,
    fill = "grey50",
    alpha = 0.4,
    position = position_nudge(x = 0.15, y = -2) # Offset shadow
  ) +
  # Main bar layer
  geom_bar(
    aes(x = reorder(country, -number_of_billionaires), y = number_of_billionaires, fill = number_of_billionaires),
    stat = "identity",
    width = 0.7,
    color = "black",
    alpha = 0.9
  ) +
  # Data labels
  geom_text(
    aes(x = reorder(country, -number_of_billionaires), y = number_of_billionaires, label = number_of_billionaires),
    vjust = -0.5,
    color = "black",
    size = 5
  ) +
  # Gradient fill for the bars
  scale_fill_gradient(low = "pink", high = "purple", name = "Billionaires") +
  # Labels and theme adjustments
  labs(
    title = "Top 10 Countries by Number of Billionaires",
    subtitle = "Data represents the countries with the highest billionaire counts",
    x = "Country",
    y = "Number of Billionaires"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",  # Hide the legend
    panel.grid.major.x = element_blank() # Remove vertical gridlines
  ) +
  coord_cartesian(clip = "off") # Prevent clipping of text outside plot

# Display the plot
plot_top_countries_billionaires



```

```{r}
# 2. Top 10 Countries by Highest Total Final Worth
top_countries_final_worth <- new_billionaires_clean %>%
  arrange(desc(total_final_worth)) %>%
  slice(1:10)

plot_top_countries_final_worth <- ggplot(top_countries_final_worth) +
  # Shadow layer (slightly offset and darker)
  geom_bar(
    aes(x = reorder(country, -total_final_worth), y = total_final_worth),
    stat = "identity",
    width = 0.7,
    fill = "grey50",
    alpha = 0.4,
    position = position_nudge(x = 0.15, y = -2) # Offset shadow
  ) +
  # Main bar layer
  geom_bar(
    aes(x = reorder(country, -total_final_worth), y = total_final_worth, fill = total_final_worth),
    stat = "identity",
    width = 0.7,
    color = "black",
    alpha = 0.9
  ) +
  # Data labels
  geom_text(
    aes(x = reorder(country, -total_final_worth), y = total_final_worth, label = scales::comma(total_final_worth)),
    vjust = -0.5,
    color = "black",
    size = 5
  ) +
  # Gradient fill for the bars
  scale_fill_gradient(low = "pink", high = "purple", name = "Total Final Worth") +
  # Labels and theme adjustments
  labs(
    title = "Top 10 Countries by Total Final Worth",
    subtitle = "Data represents the countries with the highest combined billionaire wealth",
    x = "Country",
    y = "Total Final Worth (in billions, USD)"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid.major.x = element_blank() # Remove vertical gridlines
  ) +
  coord_cartesian(clip = "off") # Prevent clipping of text outside plot

# Display the plot
plot_top_countries_final_worth


```

```{r}
# 3. Self-Made vs. Inherited Wealth (Using Proportion)
self_made_proportions <- new_billionaires_clean %>%
  summarise(
    proportion_self_made = mean(proportion_self_made, na.rm = TRUE), # Proportion of self-made
    proportion_inherited = 1 - mean(proportion_self_made, na.rm = TRUE) # Proportion of inherited
  )

# Transform data into long format for plotting
self_made_proportions_long <- self_made_proportions %>%
  pivot_longer(cols = everything(), names_to = "wealth_status", values_to = "proportion") %>%
  mutate(wealth_status = recode(wealth_status, 
                                "proportion_self_made" = "Self-Made", 
                                "proportion_inherited" = "Inherited"))

plot_self_made_vs_inherited <- ggplot(self_made_proportions_long) +
  # Shadow layer (slightly offset and darker)
  geom_bar(
    aes(x = wealth_status, y = proportion),
    stat = "identity",
    width = 0.7,
    fill = "grey50",
    alpha = 0.4,
    position = position_nudge(x = 0.15, y = -0.02) # Offset shadow
  ) +
  # Main bar layer
  geom_bar(
    aes(x = wealth_status, y = proportion, fill = wealth_status),
    stat = "identity",
    width = 0.7,
    color = "black",
    alpha = 0.9
  ) +
  # Data labels
  geom_text(
    aes(x = wealth_status, y = proportion, label = scales::percent(proportion)),
    vjust = -0.5,
    color = "black",
    size = 5
  ) +
  # Custom fill colors for Self-Made and Inherited
  scale_fill_manual(values = c("Self-Made" = "pink", "Inherited" = "purple")) +
  # Labels and theme adjustments
  labs(
    title = "Self-Made vs. Inherited Wealth Proportions",
    subtitle = "Proportions of self-made vs. inherited wealth among billionaires",
    x = "Wealth Status",
    y = "Proportion"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    legend.position = "none",  # Hide the legend
    panel.grid.major.x = element_blank() # Remove vertical gridlines
  ) +
  coord_cartesian(clip = "off") # Prevent clipping of text outside plot

# Display the plot
plot_self_made_vs_inherited

```

```{r}
# 4. Gender Distribution (Using Proportion)
gender_proportions <- new_billionaires_clean %>%
  summarise(
    proportion_male = mean(proportion_male, na.rm = TRUE),
    proportion_female = 1 - mean(proportion_male, na.rm = TRUE) # Calculate female proportion
  )

# Transform data into long format for plotting
gender_proportions_long <- gender_proportions %>%
  pivot_longer(cols = everything(), names_to = "gender", values_to = "proportion") %>%
  mutate(gender = recode(gender, "proportion_male" = "Male", "proportion_female" = "Female"))

# Add percentage labels for the pie chart
gender_proportions_long <- gender_proportions_long %>%
  mutate(label = paste0(round(proportion * 100, 1), "%"))

# Create a more special gender distribution pie chart
plot_gender <- ggplot(gender_proportions_long, aes(x = 2, y = proportion, fill = gender)) +
  geom_bar(stat = "identity", width = 1, color = "white") +  # Add border to slices
  coord_polar("y", start = 0) +
  geom_text(aes(y = cumsum(proportion) - (proportion / 2), label = label), color = "white", size = 5) +  # Labels inside slices
  labs(
    title = "Gender Distribution Across Countries",
    subtitle = "Proportion of Male and Female Billionaires"
  ) +
  scale_fill_manual(values = c("Male" = "purple", "Female" = "pink")) +  # Enhanced colors
  theme_void() +  # Cleaner theme
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  xlim(0.5, 2.5)  # Donut-style chart

plot_gender

```

```{r}
# Split concatenated industries into individual rows
industry_counts <- new_billionaires_clean %>%
  separate_rows(concatenated_industries, sep = ", ") %>%  # Split industries by ", "
  count(concatenated_industries, sort = TRUE)            # Count occurrences of each industry

# Extract the top 10 industries by count
top_industries <- industry_counts %>%
  slice(1:10)

# Generate a gradient from pink to purple
gradient_colors <- colorRampPalette(c("pink", "purple"))(10)

# Create the plot with gradient colors and counts displayed
plot_top_industries <- ggplot(top_industries, aes(x = reorder(concatenated_industries, n), y = n, fill = reorder(concatenated_industries, n))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +  # Add count labels to bars
  coord_flip() +
  scale_fill_manual(values = gradient_colors) +  # Apply gradient colors in correct order
  labs(
    title = "Top 10 Industries by Number of Billionaires",
    x = "Industry",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14)
  )

plot_top_industries

```

```{r}
# Enhanced histogram for number_of_billionaires
ggplot(new_billionaires_clean, aes(x = number_of_billionaires)) +
  geom_histogram(
    binwidth = 10, # Adjust bin width
    fill = "purple", # Updated fill color to a gradient shade
    color = "white", # Bar border for a cleaner look
    alpha = 0.85 # Slight transparency for better layering
  ) +
  scale_x_continuous(labels = scales::comma) + # Format x-axis labels for better readability
  labs(
    title = "Distribution of Billionaires Across Countries",
    subtitle = "Grouped by Number of Billionaires in Intervals of 10",
    x = "Number of Billionaires",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 15) + # Larger font size for better readability
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5), # Center and bold title
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5), # Styled subtitle
    axis.title.x = element_text(size = 14, margin = margin(t = 10)), # Add margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)), # Add margin to y-axis title
    axis.text = element_text(size = 12), # Adjust axis text size
    panel.grid.major = element_line(color = "gray85", linetype = "dashed") # Dashed grid lines for clarity
  )

```

```{r}
ggplot(new_billionaires_clean, aes(y = number_of_billionaires, x = total_final_worth)) +
  geom_point() +
  xlab("total_final_worth") +
  ylab("number_of_billionaires")
```

```{r}
# library(ggplot2)
# library(gganimate)
# library(dplyr)
# 
# # Prepare data for animation
# new_billionaires_plot <- new_billionaires_clean %>%
#   arrange(total_final_worth) %>%
#   mutate(increment = 1:nrow(new_billionaires_clean)) # Create an increment column to control the animation
# 
# # Animated scatter plot
# animated_scatter <- ggplot(new_billionaires_plot, aes(x = total_final_worth, y = number_of_billionaires)) +
#   geom_point(aes(color = "purple"), alpha = 0.7, size = 3) +  # Stylish points with transparency
#   xlab("Total Final Worth (in billions)") +
#   ylab("Number of Billionaires") +
#   scale_x_continuous(labels = scales::comma) +
#   scale_y_continuous(labels = scales::comma) +
#   labs(
#     title = "Number of Billionaires vs. Total Final Worth",
#     subtitle = "Increment: {closest_state}",
#     caption = "Data Source: Billionaires Dataset"
#   ) +
#   theme_minimal(base_size = 15) +
#   theme(
#     plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
#     plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
#     axis.title.x = element_text(size = 14),
#     axis.title.y = element_text(size = 14)
#   ) +
#   transition_states(increment, transition_length = 1, state_length = 1) +  # Reveal points gradually and keep them
#   enter_fade() +  # Fade in new points smoothly
#   shadow_mark(alpha = 1, size = 3) +  # Keep all previous points visible
#   ease_aes('linear')  # Smooth animation
# 
# # Render animation with 2-second pause at the end
# animate(
#   animated_scatter,
#   nframes = 150,          # Total frames
#   fps = 40,               # Frames per second
#   width = 800,            # Width of output
#   height = 600,           # Height of output
#   renderer = gifski_renderer(),
#   end_pause = 30          # Pause for 2 seconds (20 frames at 10 fps)
# )

```

```{r}
ggplot(new_billionaires_clean, aes(y = number_of_billionaires, x = total_final_worth)) +
  geom_point() +
  coord_cartesian(ylim = c(0, 100), xlim = c(0, 5e5)) +  # Adjust ranges as needed
  xlab("total_final_worth") +
  ylab("number_of_billionaires")
```

```{r}
# library(ggplot2)
# library(gganimate)
# library(dplyr)
# 
# # Prepare data for animation
# new_billionaires_plot <- new_billionaires_clean %>%
#   arrange(total_final_worth) %>%
#   mutate(increment = 1:nrow(new_billionaires_clean)) # Create an increment column to control the animation
# 
# # Animated scatter plot with customized limits
# animated_scatter_limited <- ggplot(new_billionaires_plot, aes(x = total_final_worth , y = number_of_billionaires)) +
#   geom_point(aes(color = "purple"), alpha = 0.7, size = 3) +  # Increase point size for better visibility
#   coord_cartesian(xlim = c(0, 5e5), ylim = c(0, 100)) +  # Adjust ranges as needed
#   xlab("Total Final Worth (in billions)") +
#   ylab("Number of Billionaires") +
#   scale_x_continuous(labels = scales::comma) +
#   scale_y_continuous(labels = scales::comma) +
#   labs(
#     title = "Number of Billionaires vs. Total Final Worth",
#     subtitle = "Increment: {closest_state}",
#     caption = "Data Source: Billionaires Dataset"
#   ) +
#   theme_minimal(base_size = 15) +
#   theme(
#     plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
#     plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
#     axis.title.x = element_text(size = 14),
#     axis.title.y = element_text(size = 14)
#   ) +
#   transition_states(increment, transition_length = 1, state_length = 1) +  # Reveal points gradually and keep them
#   enter_fade() +  # Fade in new points smoothly
#   shadow_mark(alpha = 1, size = 3) +  # Keep all previous points visible
#   ease_aes('linear')  # Smooth animation
# 
# # Render animation with 2-second pause at the end
# animate(
#   animated_scatter_limited,
#   nframes = 150,          # Total frames
#   fps = 40,               # Frames per second
#   width = 800,            # Width of output
#   height = 600,           # Height of output
#   renderer = gifski_renderer(),
#   end_pause = 30          # Pause for 2 seconds (20 frames at 10 fps)
# )


```

```{r}
iqr_value <- IQR(new_billionaires_clean$number_of_billionaires)
cat("IQR value: ",iqr_value, "\n")

# Calculate Q1 and Q3
q1 <- quantile(new_billionaires_clean$number_of_billionaires,0.25)
q3 <- quantile(new_billionaires_clean$number_of_billionaires,0.75)
cat("The Q1 is: ", q1, " and Q3 is: ", q3, '\n' )

#Identify outliers
lower_bound <- q1 - 1.5 * iqr_value
upper_bound <- q3 + 1.5 * iqr_value
lower_bound
upper_bound
outliers <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$number_of_billionaires < lower_bound | new_billionaires_clean$number_of_billionaires > upper_bound]

cat("Outliers: ", outliers)

num_outliers <- length(outliers)
cat("\nNumber of outliers: ", num_outliers, "\n")
```

```{r}
n <- length(new_billionaires_clean$number_of_billionaires)

#percentage of outliers in the dataset
percentage_o <- (num_outliers/n) *100
percentage_o
```

```{r}
# Remove outliers
new_billionaires_clean <- new_billionaires_clean[new_billionaires_clean$number_of_billionaires >= lower_bound 
                                           & new_billionaires_clean$number_of_billionaires <= upper_bound ,]
outliers_new <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$number_of_billionaires < lower_bound | new_billionaires_clean$number_of_billionaires > upper_bound]
num_outliers_new <- length(outliers_new)
num_outliers_new
```

```{r}
View(new_billionaires_clean)
```

```{r}
# Enhanced histogram for number_of_billionaires
ggplot(new_billionaires_clean, aes(x = number_of_billionaires)) +
  geom_histogram(
    binwidth = 10, # Adjust bin width
    fill = "purple", # Updated fill color to a gradient shade
    color = "white", # Bar border for a cleaner look
    alpha = 0.85 # Slight transparency for better layering
  ) +
  scale_x_continuous(labels = scales::comma) + # Format x-axis labels for better readability
  labs(
    title = "Distribution of Billionaires Across Countries",
    subtitle = "Grouped by Number of Billionaires in Intervals of 10",
    x = "Number of Billionaires",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 15) + # Larger font size for better readability
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5), # Center and bold title
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5), # Styled subtitle
    axis.title.x = element_text(size = 14, margin = margin(t = 10)), # Add margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)), # Add margin to y-axis title
    axis.text = element_text(size = 12), # Adjust axis text size
    panel.grid.major = element_line(color = "gray85", linetype = "dashed") # Dashed grid lines for clarity
  )
```

## Calculate statistical measures for number_of_billionaires

```{r}
stats <- new_billionaires_clean %>% summarise( mean = mean(number_of_billionaires, na.rm = TRUE), median = median(number_of_billionaires, na.rm = TRUE), mode = as.numeric(names(sort(table(number_of_billionaires), decreasing = TRUE)[1])), variance = var(number_of_billionaires, na.rm = TRUE), std_dev = sd(number_of_billionaires, na.rm = TRUE) )

print(stats)
```

```{r}
library(moments)
skewness_value <- skewness(new_billionaires_clean$number_of_billionaires, na.rm = TRUE)
kurtosis_value <- kurtosis(new_billionaires_clean$number_of_billionaires, na.rm = TRUE)
cat("Skewness:", skewness_value, "\nKurtosis:", kurtosis_value, "\n")
```

```{r}
# Fit a Poisson distribution
set.seed(42)
lambda <- mean(new_billionaires_clean$number_of_billionaires) # Poisson parameter λ
lambda
# Add theoretical Poisson probabilities to the dataset
max_billionaires <- max(new_billionaires_clean$number_of_billionaires)
poisson_prob <- dpois(0:max_billionaires, lambda)

# Compare observed vs expected (Poisson)
observed_counts <- table(new_billionaires_clean$number_of_billionaires)
observed_prob <- observed_counts / sum(observed_counts)
```

```{r}
# Plot observed vs Poisson probabilities
comparison <- data.frame(
  number_of_billionaires = as.numeric(names(observed_counts)),
  observed_prob = as.vector(observed_prob),
  poisson_prob = poisson_prob[1:length(observed_counts)]
)

# Plot observed vs Poisson probabilities with enhanced styling
ggplot(comparison, aes(x = number_of_billionaires)) +
  geom_bar(aes(y = observed_prob), stat = "identity", fill = "purple", alpha = 0.85, color = "white") + # Adjusted to purple and added border
  geom_line(aes(y = poisson_prob), color = "red", size = 1) + # Retained red line
  scale_x_continuous(labels = scales::comma) + # Format x-axis labels for readability
  labs(
    title = "Observed vs. Poisson Distribution",
    x = "Number of Billionaires",
    y = "Probability"
  ) +
  theme_minimal(base_size = 15) + # Larger font size for better readability
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5), # Center and bold title
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5), # Styled subtitle
    axis.title.x = element_text(size = 14, margin = margin(t = 10)), # Add margin to x-axis title
    axis.title.y = element_text(size = 14, margin = margin(r = 10)), # Add margin to y-axis title
    axis.text = element_text(size = 12), # Adjust axis text size
    panel.grid.major = element_line(color = "gray85", linetype = "dashed") # Dashed grid lines for clarity
  )

```

## Extracting sample from population

### Random sample

```{r}
set.seed(42)

random_sample <- new_billionaires_clean %>%
  sample_n(size = 40)

# View the random sample
print(random_sample)

```

### Perform one sample Z-Test

```{r}
#Perform one sample Z-Test
mean_population<- mean(new_billionaires_clean$number_of_billionaires)
cat("Population mean: ", mean_population)

mean_sample <- mean(random_sample$number_of_billionaires)
cat("\nSample mean: ", mean_sample)

sd_population <- sqrt(var(new_billionaires_clean$number_of_billionaires))

n <- length(new_billionaires_clean$number_of_billionaires)  

standard_error <- sd_population / sqrt(n)

# Calculate Z statistics
z_statistics <- (mean_sample - mean_population) / standard_error
cat("\nZ statistic: ", round(z_statistics, 2))

# Determine critical z-value for two tailed test at 0.05 significance level
alpha <- 0.05
z_critical <- qnorm(1 - alpha / 2) #two tailed use alpha / 2
cat("\nCritical value: ", round(z_critical, 2))

# Compare z statistics and critical value to make decision
if (abs(z_statistics) > z_critical) {
  decision <- "Reject the null hypothesis"
}else {
  decision <- "Fail to reject the null hypothesis"
}

cat("\nDecision of Z test:",decision)
```

### Sampling Distribution for Number of Billionaires

```{r}
# Sampling DIstribution for number of billionaires
# Parameters for simulation
sample_size <- 40  # Define the size of each sample
n_simulations <- 1000  # Number of simulations

# Simulate sample means
sample_means <- replicate(
  n_simulations,
  mean(sample(new_billionaires_clean$number_of_billionaires, size = sample_size, replace = TRUE))
)

# Summary of the sampling distribution
summary(sample_means)
```

```{r}
# Plot the sampling distribution with enhanced style
ggplot(data.frame(sample_means = sample_means), aes(x = sample_means)) +
  geom_histogram(
    binwidth = 0.5, 
    fill = "purple", 
    color = "white", 
    alpha = 0.85
  ) +
  labs(
    title = "Sampling Distribution of the Mean",
    subtitle = "Illustrating the Distribution of Sample Means",
    x = "Sample Mean",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "gray85", linetype = "dashed")
  )

theoretical_mean <- mean(sample_means)  # Should approximate the population mean
theoretical_sd <- sd(sample_means)  # Standard error of the mean

# Overlay normal curve with enhanced style
ggplot(data.frame(sample_means = sample_means), aes(x = sample_means)) +
  geom_histogram(
    aes(y = ..density..), 
    binwidth = 0.5, 
    fill = "purple", 
    color = "white", 
    alpha = 0.85
  ) +
  stat_function(
    fun = dnorm, 
    args = list(mean = theoretical_mean, sd = theoretical_sd),
    color = "red", 
    size = 1.2
  ) +
  labs(
    title = "Sampling Distribution of the Sample Mean with Normal Curve",
    subtitle = "Comparison Between Observed Sampling Distribution and Theoretical Normal Curve",
    x = "Sample Mean",
    y = "Density"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, face = "italic", hjust = 0.5),
    axis.title.x = element_text(size = 14, margin = margin(t = 10)),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "gray85", linetype = "dashed")
  )

```

# Section 1: Economic Indicators and Billionaires Distribution

## 1.1 Descriptive Statistics

### Calculate Summary Statistics

```{r}
# List of economic indicators to analyze
indicators <- c("gdp_country", "cpi_country", "cpi_change_country", 
                "tax_revenue_country", "total_tax_rate_country", "population_country")

# Summary table for mean, median, and variance
summary_stats <- new_billionaires_clean %>%
  summarise(across(all_of(indicators), 
                   list(mean = ~ mean(. , na.rm = TRUE),
                        median = ~ median(. , na.rm = TRUE),
                        variance = ~ var(. , na.rm = TRUE)
                   ), .names = "{col}_{fn}")) 

# Print the summary table
print(summary_stats)
```

```{r}
# Load required libraries
library(ggplot2)
library(patchwork)

# Create individual histograms for each indicator
hist_gdp <- ggplot(new_billionaires_clean, aes(x = gdp_country)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "GDP (Country)", x = "GDP", y = "Frequency") +
  theme_minimal()

hist_cpi <- ggplot(new_billionaires_clean, aes(x = cpi_country)) +
  geom_histogram(fill = "lightgreen", color = "black", bins = 30) +
  labs(title = "CPI (Country)", x = "CPI", y = "Frequency") +
  theme_minimal()

hist_cpi_change <- ggplot(new_billionaires_clean, aes(x = cpi_change_country)) +
  geom_histogram(fill = "orange", color = "black", bins = 30) +
  labs(title = "CPI Change (Country)", x = "CPI Change", y = "Frequency") +
  theme_minimal()

hist_tax_revenue <- ggplot(new_billionaires_clean, aes(x = tax_revenue_country)) +
  geom_histogram(fill = "purple", color = "black", bins = 30) +
  labs(title = "Tax Revenue (Country)", x = "Tax Revenue (%)", y = "Frequency") +
  theme_minimal()

hist_tax_rate <- ggplot(new_billionaires_clean, aes(x = total_tax_rate_country)) +
  geom_histogram(fill = "pink", color = "black", bins = 30) +
  labs(title = "Total Tax Rate (Country)", x = "Tax Rate (%)", y = "Frequency") +
  theme_minimal()


# Combine all histograms in one subplot using patchwork
(hist_gdp | hist_cpi | hist_cpi_change) /
  (hist_tax_revenue | hist_tax_rate )
```

## 1.2 Correlation Analysis

```{r}
# Extract data only for the selected indicators

indicators <- c("number_of_billionaires", "gdp_country", "cpi_country", "cpi_change_country", 
                "tax_revenue_country", "total_tax_rate_country", "population_country")
data1 <- new_billionaires_clean[indicators]
```

### Covariance Matrix

```{r}
cov_matrix <- cov(data1, use = "complete.obs")
print("Covariance Matrix:")
print(cov_matrix)
```

### Pearson Correlation

```{r}
pearson_corr <- cor(data1, method = "pearson", use = "complete.obs")
print("Pearson Correlation Matrix:")
print(pearson_corr)
```

### Spearman Correlation

```{r}
spearman_corr <- cor(data1, method = "spearman", use = "complete.obs")
print("Spearman Correlation Matrix:")
print(spearman_corr)
```

### Visualisation for Summary

```{r}
# Convert matrices to long format for ggplot2 heatmap
pearson_long <- melt(pearson_corr)
spearman_long <- melt(spearman_corr)

# Function to create a heatmap with text annotations
create_heatmap <- function(data_long, title) {
  ggplot(data_long, aes(Var1, Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3) +
    scale_fill_gradient2(low = "pink", high = "purple", mid = "white", 
                         midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation") +
    labs(title = title, x = "", y = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
# Pearson Correlation Matrix Heatmap
pearson_plot <- create_heatmap(pearson_long, "Pearson Correlation Heatmap")
print(pearson_plot)
```

```{r}
# Spearman Correlation Matrix Heatmap
spearman_plot <- create_heatmap(spearman_long, "Spearman Correlation Heatmap")
print(spearman_plot)
```

## 1.3 Economic indicator gdp_country with number of billionaires

### Sample creation and threshold on gdp_country

```{r}
median_gdp <- median(new_billionaires_clean$gdp_country, na.rm = TRUE)
new_billionaires_clean$gdp_group <- ifelse(new_billionaires_clean$gdp_country > median_gdp, "High-GDP", "Low-GDP")

high_gdp_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$gdp_group == "High-GDP"]
low_gdp_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$gdp_group == "Low-GDP"]

# Summarize country-level GDP, billionaires, and group assignment
median_gdp_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_gdp = sum(gdp_country, na.rm = TRUE),          # Total GDP for the country
    gdp_group = ifelse(country_gdp > median_gdp, "High-GDP", "Low-GDP") # Group based on median GDP
  )

# Display the table with GDP summary and groups
print(median_gdp_summary)

```

### Merge the GDP group labels (`High-GDP` or `Low-GDP`) with the `new_billionaires_clean`

```{r}

table(new_billionaires_clean$gdp_group)

```

### 1.3.1 Hypothesis Testing

```{r}
# Print the null and alternative hypotheses
cat("Null Hypothesis (H0): The mean number of billionaires in the High-GDP group is equal to the mean number of billionaires in the Low-GDP group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-GDP group is not equal to the mean number of billionaires in the Low-GDP group.\n\n")
```

```{r}
# Step 1: Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_gdp_group)
shapiro_low <- shapiro.test(low_gdp_group)

cat("Shapiro-Wilk Test for High-GDP Group:\n")
print(shapiro_high)

cat("\nShapiro-Wilk Test for Low-GDP Group:\n")
print(shapiro_low)

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ gdp_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ gdp_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ gdp_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = gdp_group, y = number_of_billionaires, fill = gdp_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-GDP" = "steelblue", "Low-GDP" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by GDP Group",
       x = "GDP Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.5, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", formatC(mw_test_result$p.value, format = "e", digits = 2), "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()
```

### 1.3.2 Regression Analysis

```{r}
# Regression Analysis
# Aim: Predict Y= Number of billionaires based on gdp_country , cpi....
# Perform Simple Linear Regression
# Example: Simple Linear Regression
model_simple <- lm(number_of_billionaires ~ gdp_country, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = gdp_country, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression", x = "GDP (Country)", y = "Number of Billionaires")

# residuals 
residuals <- resid(model_simple)
# Plot residuals
plot(new_billionaires_clean$gdp_country, residuals, main = "Residuals for Scenario 1", 
     xlab = "Size (sq ft)", ylab = "Residuals", pch = 19)
abline(h = 0, col = "red")

```

## 1.4 Economic indicator cpi_country with number of billionaires

### Sample creation and threshold on cpi_country

```{r}
# Create CPI group column based on median GDP
median_cpi <- median(new_billionaires_clean$cpi_country, na.rm = TRUE)
new_billionaires_clean$cpi_group <- ifelse(new_billionaires_clean$cpi_country > median_cpi, "High-CPI", "Low-CPI")

# Separate the data into High-CPI and Low-CPI groups
high_cpi_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$cpi_group == "High-CPI"]
low_cpi_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$cpi_group == "Low-CPI"]

median_cpi_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_cpi = sum(cpi_country, na.rm = TRUE),     
    cpi_group = ifelse(country_cpi > median_cpi, "High-cpi", "Low-cpi") 
  )

print(median_cpi_summary)
```

### Merge the CPI group labels (`High-CPI`or `Low-CPI`) with the `new_billionaires_clean`

```{r}
# Check the updated data
table(new_billionaires_clean$cpi_group)
```

### 1.4.1 Hypothesis Testing

```{r}
cat("Null Hypothesis (H0): The mean number of billionaires in the High-CPI group is equal to the mean number of billionaires in the Low-CPI group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-CPI group is not equal to the mean number of billionaires in the Low-CPI group.\n\n")

```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_cpi_group)
shapiro_low <- shapiro.test(low_cpi_group)

cat("Shapiro-Wilk Test for High-CPI Group:\n")
print(shapiro_high)

cat("\nShapiro-Wilk Test for Low-CPI Group:\n")
print(shapiro_low)

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ cpi_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ cpi_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ cpi_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}

```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = cpi_group, y = number_of_billionaires, fill = cpi_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-CPI" = "steelblue", "Low-CPI" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by CPI Group",
       x = "CPI Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.4, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", round(mw_test_result$p.value,5), "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()

```

### Regression Analysis

```{r}
# Regression Analysis
# Aim: Predict Y= Number of billionaires based on gdp_country , cpi....
# Perform Simple Linear Regression
# Example: Simple Linear Regression
model_simple <- lm(number_of_billionaires ~ cpi_country, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = cpi_country, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression CPI vs Numbers of Billionaires", x = "CPI (Country)", y = "Number of Billionaires")

```

## 1.5 Economic indicator cpi_change_country with number of billionaires

### Sample creation and threshold on cpi_change_country

```{r}
# Create CPI-change group column based on median CPI-change
median_cpi_change_country <- median(new_billionaires_clean$cpi_change_country, na.rm = TRUE)
new_billionaires_clean$cpi_change_country_group <- ifelse(new_billionaires_clean$cpi_change_country > median_cpi_change_country, "High-CPI-change", "Low-CPI-change")

# Separate the data into High-CPI and Low-CPI groups
high_cpi_change_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$cpi_change_country_group == "High-CPI-change"]
low_cpi_change_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$cpi_change_country_group == "Low-CPI-change"]

median_cpi_change_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_cpi_change = sum(cpi_change_country, na.rm = TRUE),     
    cpi_change_country_group = ifelse(country_cpi_change > median_cpi_change_country, "High-CPI_change", "Low-CPI-change") 
  )

print(median_cpi_change_summary)
```

### Merge the tax group labels (`High-cpi-change`or `Low-cpi-change`) with the `new_billionaires_clean`

```{r}
table(new_billionaires_clean$cpi_change_country_group)
```

### 1.5.1 Hypothesis Testing

```{r}
cat("Null Hypothesis (H0): The mean number of billionaires in the High-cpi-change group is equal to the mean number of billionaires in the Low-cpi-change group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-cpi-change group is not equal to the mean number of billionaires in the Low-cpi-change group.\n\n")
```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_cpi_change_group)
shapiro_low <- shapiro.test(low_cpi_change_group)
print("Shapiro-Wilk Test for High CPI-change Group:\n")
shapiro_high
print("Shapiro-Wilk Test for Low CPI-change Group:\n")
shapiro_low

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ cpi_change_country_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ cpi_change_country_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ cpi_change_country_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = cpi_change_country_group, y = number_of_billionaires, fill = cpi_change_country_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-CPI-change" = "steelblue", "Low-CPI-change" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by CPI-change Group",
       x = "CPI-Change Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.2, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", round(mw_test_result$p.value,5) , "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()
```

### Regression Analysis

```{r}
# Regression Analysis
# Aim: Predict Y= Number of billionaires based on gdp_country , cpi....
# Perform Simple Linear Regression
# Example: Simple Linear Regression
model_simple <- lm(number_of_billionaires ~ cpi_change_country, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = cpi_change_country, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression CPI change vs Number of Billionaires", x = "CPI-change (Country)", y = "Number of Billionaires")

# residuals 
residuals <- resid(model_simple)
# Plot residuals
plot(new_billionaires_clean$cpi_change_country, residuals, main = "Residuals for Scenario 3", 
     xlab = "Size (sq ft)", ylab = "Residuals", pch = 19)
abline(h = 0, col = "red")
```

## 1.6 Economic indicator total_tax_rate_country with number of billionaires

### Sample creation and threshold on total_tax_rate_country

```{r}
# Create total tax rate group column based on median total tax rate
median_total_tax_rate_country <- median(new_billionaires_clean$total_tax_rate_country, na.rm = TRUE)
new_billionaires_clean$total_tax_rate_country_group <- ifelse(new_billionaires_clean$total_tax_rate_country > median_total_tax_rate_country, "High TTR", "Low TTR")

# Separate the data into High-CPI and Low-CPI groups
high_TTR_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$total_tax_rate_country_group == "High TTR"]
low_TTR_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$total_tax_rate_country_group == "Low TTR"]


median_total_tax_rate_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_total_tax_rate = sum(total_tax_rate_country, na.rm = TRUE),     
    total_tax_rate_country_group = ifelse(country_total_tax_rate > median_total_tax_rate_country, "High TTR", "Low TTR") 
  )

print(median_total_tax_rate_summary)
```

### Merge the tax group labels (`High TTR`or `Low TTR`) with the `new_billionaires_clean`

```{r}
# Check the updated data
table(new_billionaires_clean$total_tax_rate_country_group)
```

### 1.5.1 Hypothesis Testing

```{r}
cat("Null Hypothesis (H0): The mean number of billionaires in the High TTR group is equal to the mean number of billionaires in the Low TTR group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High TTR group is not equal to the mean number of billionaires in the Low TTR group.\n\n")

```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_TTR_group)
shapiro_low <- shapiro.test(low_TTR_group)
print("Shapiro-Wilk Test for High Total Tax Rate Group:\n")
shapiro_high
print("Shapiro-Wilk Test for Low Total Tax Rate Group:\n")
shapiro_low

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ total_tax_rate_country_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ total_tax_rate_country_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ total_tax_rate_country_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = total_tax_rate_country_group, y = number_of_billionaires, fill = total_tax_rate_country_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High TTR" = "steelblue", "Low TTR" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by Total Tax Rate (TTR) Group",
       x = "TTR Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 2, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", round(mw_test_result$p.value,5) , "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()
```

# Section 2: Socioeconomic Indicators on Billionaire Distribution

## 2.1 Descriptive Statistics

### Calculate Summary Statistics

```{r}
# List of socioeconomic indicators to analyze
socioeconomic_indicators <- c("population_country", "life_expectancy_country")

# Summary table for mean, median, and variance
summary_stats <- new_billionaires_clean %>%
  summarise(across(all_of(socioeconomic_indicators), 
                   list(mean = ~ mean(. , na.rm = TRUE),
                        median = ~ median(. , na.rm = TRUE),
                        variance = ~ var(. , na.rm = TRUE)
                   ), .names = "{col}_{fn}")) 

# Print the summary table
print(summary_stats)
```

```{r}
# Load required libraries
library(ggplot2)
library(patchwork)

hist_population <- ggplot(new_billionaires_clean, aes(x = population_country)) +
  geom_histogram(fill = "purple", color = "black", bins = 30) +
  labs(title = "Population (Country)", x = "Population", y = "Frequency") +
  theme_minimal()

hist_life_expectancy <- ggplot(new_billionaires_clean, aes(x = life_expectancy_country)) +
  geom_histogram(fill = "pink", color = "black", bins = 30) +
  labs(title = "Life Expectancy (Country)", x = "Life Expectancy (%)", y = "Frequency") +
  theme_minimal()


# Combine all histograms in one subplot using patchwork
(hist_population| hist_life_expectancy)
```

## 2.2 Correlation Analysis

```{r}
# Extract data only for the selected indicators

socioeconomic_indicators <- c("number_of_billionaires", "population_country", "life_expectancy_country")
data2 <- new_billionaires_clean[socioeconomic_indicators]
```

### Covariance Matrix

```{r}
cov_matrix <- cov(data2, use = "complete.obs")
print("Covariance Matrix:")
print(cov_matrix)
```

### Pearson Correlation

```{r}
pearson_corr <- cor(data2, method = "pearson", use = "complete.obs")
print("Pearson Correlation Matrix:")
print(pearson_corr)
```

### Spearman Correlation

```{r}
spearman_corr <- cor(data2, method = "spearman", use = "complete.obs")
print("Spearman Correlation Matrix:")
print(spearman_corr)
```

### Visualisation for Summary

```{r}
# Convert matrices to long format for ggplot2 heatmap
pearson_long <- melt(pearson_corr)
spearman_long <- melt(spearman_corr)

# Function to create a heatmap with text annotations
create_heatmap <- function(data_long, title) {
  ggplot(data_long, aes(Var1, Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3) +
    scale_fill_gradient2(low = "pink", high = "purple", mid = "white", 
                         midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation") +
    labs(title = title, x = "", y = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
# Pearson Correlation Matrix Heatmap
pearson_plot <- create_heatmap(pearson_long, "Pearson Correlation Heatmap")
print(pearson_plot)
```

```{r}
# Spearman Correlation Matrix Heatmap
spearman_plot <- create_heatmap(spearman_long, "Spearman Correlation Heatmap")
print(spearman_plot)
```

## 2.3 Socioeconomic indicator population_country with number of billionaires

### Sample creation and threshold on population

```{r}
median_population <- median(new_billionaires_clean$population_country, na.rm = TRUE)
new_billionaires_clean$population_group <- ifelse(new_billionaires_clean$population_country > median_population, "High-Population", "Low-Population")

high_population_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$population_group == "High-Population"]
low_population_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$population_group == "Low-Population"]

median_population_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_population = sum(population_country, na.rm = TRUE),  
    population_group = ifelse(country_population > median_population, "High-Population", "Low-Population") 
  )


print(median_population_summary)
```

### Merge the tax group labels (`High-population`or `Low-population`) with the `new_billionaires_clean`

```{r}
table(new_billionaires_clean$population_group)
```

### 2.3.1 Hypothesis Testing

```{r}
# Print the null and alternative hypotheses
cat("Null Hypothesis (H0): The mean number of billionaires in the High-Population group is equal to the mean number of billionaires in the Low-Population group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-Population group is not equal to the mean number of billionaires in the Low-Population group.\n\n")
```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_population_group)
shapiro_low <- shapiro.test(low_population_group)

cat("\nShapiro-Wilk Test for High-Population Group:\n")
print(shapiro_high)

cat("\nShapiro-Wilk Test for Low-Population Group:\n")
print(shapiro_low)

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nResult: Both groups are normally distributed.\n")
} else {
  cat("\nResult: One or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ population_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ population_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ population_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}

```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = population_group, y = number_of_billionaires, fill = population_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-Population" = "steelblue", "Low-Population" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by Population Group",
       x = "Population Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.5, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", formatC(mw_test_result$p.value, format = "e", digits = 2), "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()

```

## 2.4 Socioeconomic indicator life_expectancy_country with number of billionaires

### Sample creation and threshold on life_expectancy_country

```{r}
median_life <- median(new_billionaires_clean$life_expectancy_country, na.rm = TRUE)
new_billionaires_clean$life_group <- ifelse(new_billionaires_clean$life_expectancy_country > median_life, "High-Life Expectancy", "Low-Life Expectancy")

high_life_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$life_group == "High-Life Expectancy"]
low_life_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$life_group == "Low-Life Expectancy"]

median_life_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_life = sum(life_expectancy_country, na.rm = TRUE),  
    life_group = ifelse(country_life > median_life, "High-Life Expectancy", "Low-Life Expectancy") 
  )


print(median_life_summary)

```

### Merge the tax group labels (`High-life expectancy`or `Low-life expectancy`) with the `new_billionaries_clean`

```{r}
table(new_billionaires_clean$life_group)
```

### 2.4.1 Hypothesis Testing

```{r}
# Print the null and alternative hypotheses
cat("Null Hypothesis (H0): The mean number of billionaires in the High-Life-Expectancy group is equal to the mean number of billionaires in the Low-Life-Expectancy group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-Life-Expectancy group is not equal to the mean number of billionaires in the Low-Life-Expectancy group.\n\n")
```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_life_group)
shapiro_low <- shapiro.test(low_life_group)

cat("\nShapiro-Wilk Test for High-Life-Expectancy Group:\n")
print(shapiro_high)

cat("\nShapiro-Wilk Test for Low-Life-Expectancy Group:\n")
print(shapiro_low)

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nResult: Both groups are normally distributed.\n")
} else {
  cat("\nResult: One or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ life_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ life_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ life_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = life_group, y = number_of_billionaires, fill = life_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-Life Expectancy" = "steelblue", "Low-Life Expectancy" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by Life Expectancy Group",
       x = "Life Expectancy Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.5, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", formatC(mw_test_result$p.value, format = "e", digits = 2), "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()

```

## 2.4.2 Regression Analysis

```{r}
# Regression Analysis
# Example: Simple Linear Regression
model_simple <- lm(number_of_billionaires ~ life_expectancy_country, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = life_expectancy_country, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression", x = "Life Expectancy (Country)", y = "Number of Billionaires")

# residuals 
residuals <- resid(model_simple)
# Plot residuals
plot(new_billionaires_clean$life_expectancy_country, residuals, main = "Plot of Residuals", 
     xlab = "yhat", ylab = "Residuals", pch = 19)
abline(h = 0, col = "red")
```

# Section 3: Industry Indicators on Billionaire Distribution

## 3.1 Descriptive Statistics

### Calculate Summary Statistics

```{r}
# List of socioeconomic indicators to analyze
industry_indicators <- c("industry_diversity")

# Summary table for mean, median, and variance
summary_stats <- new_billionaires_clean %>%
  summarise(across(all_of(industry_indicators), 
                   list(mean = ~ mean(. , na.rm = TRUE),
                        median = ~ median(. , na.rm = TRUE),
                        variance = ~ var(. , na.rm = TRUE)
                   ), .names = "{col}_{fn}")) 

# Print the summary table
print(summary_stats)

```

```{r}
hist_industry <- ggplot(new_billionaires_clean, aes(x = industry_diversity)) +
  geom_histogram(fill = "purple", color = "black", bins = 30) +
  labs(title = "Industry Types (Country)", x = "Industry Types", y = "Frequency") +
  theme_minimal()

# Combine all histograms in one subplot using patchwork
(hist_industry)
```

```{r}
industry_distribution <- new_billionaires_clean %>%
  separate_rows(concatenated_industries, sep = ",") %>%  # Split concatenated industries
  group_by(concatenated_industries) %>%
  summarise(total_billionaires = sum(number_of_billionaires, na.rm = TRUE)) %>%
  arrange(desc(total_billionaires))

# Create the treemap
treemap(
  industry_distribution,
  index = "concatenated_industries",   # The industry names
  vSize = "total_billionaires",       # The number of billionaires as block size
  title = "Distribution of Billionaire by Industry",
  palette = "Set3"                    # Use a color palette for differentiation
)
```

## 3.2 Correlation Analysis

```{r}
# Extract data only for the selected indicators

industry_indicators <- c("number_of_billionaires", "industry_diversity")
data3 <- new_billionaires_clean[industry_indicators]
```

### Covariance Matrix

```{r}
# Compute the covariance matrix
cov_matrix <- cov(data3, use = "complete.obs")

# Print the covariance matrix
print("Covariance Matrix:")
print(cov_matrix)
```

### Pearson Correlation

```{r}
pearson_corr <- cor(data3, method = "pearson", use = "complete.obs")
print("Pearson Correlation Matrix:")
print(pearson_corr)
```

### Spearman Correlation

```{r}
spearman_corr <- cor(data3, method = "spearman", use = "complete.obs")
print("Spearman Correlation Matrix:")
print(spearman_corr)
```

### Visualisation for Summary

```{r}
# Convert matrices to long format for ggplot2 heatmap
pearson_long <- melt(pearson_corr)
spearman_long <- melt(spearman_corr)

# Function to create a heatmap with text annotations
create_heatmap <- function(data_long, title) {
  ggplot(data_long, aes(Var1, Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3) +
    scale_fill_gradient2(low = "pink", high = "purple", mid = "white", 
                         midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation") +
    labs(title = title, x = "", y = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
# Pearson Correlation Matrix Heatmap
pearson_plot <- create_heatmap(pearson_long, "Pearson Correlation Heatmap")
print(pearson_plot)
```

```{r}
# Spearman Correlation Matrix Heatmap
spearman_plot <- create_heatmap(spearman_long, "Spearman Correlation Heatmap")
print(spearman_plot)
```

## 3.3 Industry indicator industry_diversity with number of billionaires

### Sample creation and threshold on gdp_country

```{r}
median_diversity <- median(new_billionaires_clean$industry_diversity, na.rm = TRUE)
new_billionaires_clean$diversity_group <- ifelse(new_billionaires_clean$industry_diversity > median_diversity, "High-Diversity", "Low-Diversity")

high_diversity_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$diversity_group == "High-Diversity"]
low_diversity_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$diversity_group == "Low-Diversity"]

median_diversity_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    diversity_industry = sum(industry_diversity, na.rm = TRUE),  
    diversity_group = ifelse(diversity_industry > median_diversity, "High-Diversity", "Low-Diversity") 
  )


print(median_diversity_summary)
```

### Merge the Diversity group labels (`High-Diversity` or `Low-Diversity`) with the `new_billionaires_clean`

```{r}
table(new_billionaires_clean$diversity_group)
```

### 3.3.1 Hypothesis Testing

```{r}
# Print the null and alternative hypotheses
cat("Null Hypothesis (H0): The mean number of billionaires in the High-Diversity group is equal to the mean number of billionaires in the Low-Diversity group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-Diversity group is not equal to the mean number of billionaires in the Low-Diversity group.\n\n")
```

```{r}
# Step 1: Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_diversity_group)
shapiro_low <- shapiro.test(low_diversity_group)

cat("Shapiro-Wilk Test for High-Diversity Group:\n")
print(shapiro_high)

cat("\nShapiro-Wilk Test for Low-Diversity Group:\n")
print(shapiro_low)

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    test_result <- t.test(
      number_of_billionaires ~ diversity_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    test_result <- t.test(
      number_of_billionaires ~ diversity_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  test_result <- wilcox.test(
    number_of_billionaires ~ diversity_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(test_result)
  
  # Decision for Mann-Whitney U Test
  if (test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Extract the p-value and decision from the t-test result
p_value <- round(test_result$p.value, 6) # Round the p-value to 4 decimal places
decision_text <- ifelse(decision == "Reject the null hypothesis H_0", "Reject Ho", "Fail to Reject Ho")

# Create the plot
ggplot(new_billionaires_clean, aes(x = diversity_group, y = number_of_billionaires, fill = diversity_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High-Diversity" = "steelblue", "Low-Diversity" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by Industry Diversity Group",
       x = "Industry Diversity Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.5, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", p_value, "\n", decision_text), 
           size = 4, 
           color = "red") +
  theme_minimal()
```

### 3.3.2 Regression Analysis

```{r}
# Regression Analysis
model_simple <- lm(number_of_billionaires ~ industry_diversity, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = industry_diversity, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression", x = "Industry Diversity (Country)", y = "Number of Billionaires")

# residuals 
residuals <- resid(model_simple)
# Plot residuals
plot(new_billionaires_clean$gdp_country, residuals, main = "Residuals Plot", 
     xlab = "yhat", ylab = "Residuals", pch = 19)
abline(h = 0, col = "red")
```

# Section 4: Education Indicators on Billionaire Distribution

## 4.1 Descriptive Statistics

### Calculate Summary Statistics

```{r}
education_indicators <- c("gross_tertiary_education_enrollment", "gross_primary_education_enrollment_country")

# Summary table for mean, median, and variance
summary_stats <- new_billionaires_clean %>%
  summarise(across(all_of(education_indicators), 
                   list(mean = ~ mean(. , na.rm = TRUE),
                        median = ~ median(. , na.rm = TRUE),
                        variance = ~ var(. , na.rm = TRUE)
                   ), .names = "{col}_{fn}")) 

# Print the summary table
print(summary_stats)
```

```{r}
# Create individual histograms for each indicator
hist_ter <- ggplot(new_billionaires_clean, aes(x = gross_tertiary_education_enrollment)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Gross tertiary education enrollment", x = "gross_tertiary_education_enrollment", y = "Frequency") +
  theme_minimal()

hist_pri <- ggplot(new_billionaires_clean, aes(x = gross_primary_education_enrollment_country)) +
  geom_histogram(fill = "lightgreen", color = "black", bins = 30) +
  labs(title = "Gross Primary education enrollment", x = "gross_primary_education_enrollment_country", y = "Frequency") +
  theme_minimal()

# Combine all histograms in one subplot using patchwork
(hist_ter | hist_pri ) 
```

## 4.2 Correlation Analysis

```{r}
indicators <- c("number_of_billionaires","gross_tertiary_education_enrollment", "gross_primary_education_enrollment_country")

data4 <- new_billionaires_clean[indicators]
```

### Covariance Matrix

```{r}
cov_matrix <- cov(data4, use = "complete.obs")
print("Covariance Matrix:")
print(cov_matrix)
```

### Pearson Correlation

```{r}
pearson_corr <- cor(data4, method = "pearson", use = "complete.obs")
print("Pearson Correlation Matrix:")
print(pearson_corr)
```

### Spearman Correlation

```{r}
spearman_corr <- cor(data4, method = "spearman", use = "complete.obs")
print("Spearman Correlation Matrix:")
print(spearman_corr)
```

### Visualisation for Summary

```{r}
# Convert matrices to long format for ggplot2 heatmap
pearson_long <- melt(pearson_corr)
spearman_long <- melt(spearman_corr)

# Function to create a heatmap with text annotations
create_heatmap <- function(data_long, title) {
  ggplot(data_long, aes(Var1, Var2, fill = value)) +
    geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3) +
    scale_fill_gradient2(low = "pink", high = "purple", mid = "white", 
                         midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation") +
    labs(title = title, x = "", y = "") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
# Pearson Correlation Matrix Heatmap
pearson_plot <- create_heatmap(pearson_long, "Pearson Correlation Heatmap")
print(pearson_plot)
```

```{r}
# Spearman Correlation Matrix Heatmap
spearman_plot <- create_heatmap(spearman_long, "Spearman Correlation Heatmap")
print(spearman_plot)
```

## 4.3 Education indicator gross_tertiary_education_enrollment with number of billionaires

### Sample creation and threshold on gross_tertiary_education

```{r}
# Create tertiary enrollment group column based on median tertiary enrollment
median_ter_enroll <- median(new_billionaires_clean$gross_tertiary_education_enrollment, na.rm = TRUE)
new_billionaires_clean$ter_enroll_group <- ifelse(new_billionaires_clean$gross_tertiary_education_enrollment > median_ter_enroll, "High_Tertiary_Enrollment", "Low_Tertiary_Enrollment")

# Separate the data into High-CPI and Low-CPI groups
high_ter_enroll_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$ter_enroll_group == "High_Tertiary_Enrollment"]
low_ter_enroll_group <- new_billionaires_clean$number_of_billionaires[new_billionaires_clean$ter_enroll_group == "Low_Tertiary_Enrollment"]

median_ter_enroll_summary <- new_billionaires_clean %>%
  group_by(country) %>%
  summarise(
    country_gross_tertiary_education_enrollment = sum(gross_tertiary_education_enrollment, na.rm = TRUE),     
    ter_enroll_group = ifelse(country_gross_tertiary_education_enrollment > median_ter_enroll, "High_Tertiary_Enrollment", "Low_Tertiary_Enrollment") 
  )

print(median_ter_enroll_summary)
```

### Merge the education group labels (`High-tertiary enrollment` or `Low-tertiary enrollment`) with the `new_billionaires_clean`

```{r}
table(new_billionaires_clean$ter_enroll_group)
```

### 4.4.1 Hypothesis Testing

```{r}
# Print the null and alternative hypotheses
cat("Null Hypothesis (H0): The mean number of billionaires in the High-Tertiary-Enrollment group is equal to the mean number of billionaires in the Low-Tertiary-Enrollment group.\n")
cat("\nAlternative Hypothesis (Ha): The mean number of billionaires in the High-Tertiary-Enrollment group is not equal to the mean number of billionaires in the Low-Tertiary-Enrollment group.\n\n")
```

```{r}
# Shapiro-Wilk Test for Normality
shapiro_high <- shapiro.test(high_ter_enroll_group)
shapiro_low <- shapiro.test(low_ter_enroll_group)
cat("Shapiro-Wilk Test for High Tertiaty Education Enrollment Group: \n")
shapiro_high
cat("Shapiro-Wilk Test for Low Tertiaty Education Enrollment Group: \n")
shapiro_low

if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed.\n")
} else {
  cat("\nOne or both groups are not normally distributed.\n")
}
```

```{r}
# Perform appropriate test based on Shapiro-Wilk p-values and F-test
if (shapiro_high$p.value > 0.05 & shapiro_low$p.value > 0.05) {
  cat("\nBoth groups are normally distributed. Proceeding with t-test.\n")
  
  # Perform t-test based on F-test result
  if (f_test$p.value > 0.05) {
    t_test_result <- t.test(
      number_of_billionaires ~ ter_enroll_group,
      data = new_billionaires_clean,
      var.equal = TRUE
    )
    cat("\nUsing independent t-test (equal variances):\n")
  } else {
    t_test_result <- t.test(
      number_of_billionaires ~ ter_enroll_group,
      data = new_billionaires_clean,
      var.equal = FALSE
    )
    cat("\nUsing Welch's t-test (unequal variances):\n")
  }
  print(t_test_result)
  
  # Critical t-Value and Decision
  alpha <- 0.05
  df <- t_test_result$parameter
  t_critical <- qt(1 - alpha / 2, df) # Two-tailed test critical value
  
  cat("\nCalculated critical t-value: ", round(t_critical, 4), "\n")
  
  # Decision
  if (abs(t_test_result$statistic) > t_critical) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
  
} else {
  cat("\nOne or both groups are not normally distributed. Proceeding with Mann-Whitney U test:\n")
  
  # Perform Mann-Whitney U Test (Wilcoxon Rank-Sum Test)
  mw_test_result <- wilcox.test(
    number_of_billionaires ~ ter_enroll_group,
    data = new_billionaires_clean,
    exact = FALSE
  )
  print(mw_test_result)
  
  # Decision for Mann-Whitney U Test
  if (mw_test_result$p.value < 0.05) {
    decision <- "Reject the null hypothesis H_0"
  } else {
    decision <- "Fail to reject the null hypothesis H_0"
  }
  cat("\nDecision: ", decision)
}
```

```{r}
# Create the plot
ggplot(new_billionaires_clean, aes(x = ter_enroll_group, y = number_of_billionaires, fill = ter_enroll_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("High_Tertiary_Enrollment" = "steelblue", "Low_Tertiary_Enrollment" = "skyblue")) +
  labs(title = "Comparison of Number of Billionaires by Tertiary Enrollment Group",
       x = "Tertiary Enrollment Group",
       y = "Number of Billionaires") +
  annotate("text", 
           x = 1.8, 
           y = max(new_billionaires_clean$number_of_billionaires, na.rm = TRUE), 
           label = paste("p-value =", formatC(mw_test_result$p.value, format = "e", digits = 2), "\n", decision), 
           size = 4, 
           color = "red") +
  theme_minimal()
```

### Regression Analysis

```{r}
# Regression Analysis
# Aim: Predict Y= Number of billionaires based on education 
# Perform Simple Linear Regression
# Example: Simple Linear Regression
model_simple <- lm(number_of_billionaires ~ gross_tertiary_education_enrollment, data = new_billionaires_clean)
# Summarize results
summary(model_simple)
# Visualize the regression line
ggplot(new_billionaires_clean, aes(x = gross_tertiary_education_enrollment, y = number_of_billionaires)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Simple Linear Regression Gross Tertiary Education Enrollment vs Number of Billionaires", x = "Gross Tertiary Education Enrollment", y = "Number of Billionaires")

# residuals 
residuals <- resid(model_simple)
# Plot residuals
plot(new_billionaires_clean$cpi_change_country, residuals, main = "Residuals", 
     xlab = "Size (sq ft)", ylab = "Residuals", pch = 19)
abline(h = 0, col = "red")

```

# Multiple Linear Regression

```{r}
# Example: Multiple Linear Regression
library(caTools)
set.seed(42)

#Splitting the dataset into training and testing data
split = sample.split(new_billionaires_clean$number_of_billionaires, SplitRatio = 0.8)
training_set = subset(new_billionaires_clean, split == TRUE)
test_set = subset(new_billionaires_clean, split == FALSE)
  
model_multiple <- lm(number_of_billionaires ~ gdp_country + life_expectancy_country + industry_diversity,
                     data = training_set)
summary(model_multiple)
```

### Reduced Model

```{r}
model_multiple <- lm(number_of_billionaires ~ gdp_country + industry_diversity,
                     data = training_set)
summary(model_multiple)
```

### Model Diagnostics

```{r}
# 2. Model Diagnostics
# Calculate and plot the residuals
par(mfrow=c(1, 2))
plot(model_multiple$fitted.values, model_multiple$residuals, 
     main="Residuals vs Fitted", xlab="Fitted values", ylab="Residuals")
# Add a horizontal line at y = 0 to help assess residuals
abline(h=0, col="red")

```

```{r}
#Plot the Cook Distance
cooksD <- cooks.distance(model_multiple) 
plot(cooksD, main="Cook's Distance", ylab="Cook's distance")
n <- nrow(training_set)
abline(h = 4/n, lty = 2, col = "red") # add cutoff line

#Identify influential (outlier) data points
influential_obs <- as.numeric(names(cooksD)[(cooksD > (4/n))])
#define new data frame with influential points removed
outliers_removed <- training_set[-influential_obs, ]
```

### Model Diagnostics after removing outliers

```{r}
#Visualizes  residuals against the predictors before and after outliers removed.
par(mfrow = c(1, 2))
#Before
plot(model_multiple$fitted.values,model_multiple$residuals, 
     main="Before: Residuals vs Fitted", xlab="Fitted values", ylab="Residuals")
abline(h = 0, col = "red", lty = 2)

```

```{r}
#After
model_after <- lm(number_of_billionaires ~ gdp_country + life_expectancy_country + industry_diversity,
                  data = outliers_removed)
plot(model_after$fitted.values, model_after$residuals, 
     main="After: Residuals vs Fitted", xlab="Fitted values", ylab="Residuals")
abline(h = 0, col = "red", lty = 2)

```

```{r}
# Check for multicollinearity using variance Inflation Factor (VIF)
library(car)
vif_values <- vif(model_multiple)
print(vif_values)
```

### Predictions

```{r}
# Use the model to predict number of billionaires and the Confidence Interval for the new data
predicted_billionaires <- predict(model_multiple, test_set, 
                                  interval = "confidence", level = 0.95)
predicted_billionaires

```

```{r}
# Combine actual and predicted values into the test set
test_set_with_predictions <- cbind(test_set, predicted_billionaires)

# Rename prediction columns for clarity
names(test_set_with_predictions)[names(test_set_with_predictions) == "fit"] <- "Predicted_Billionaires"
names(test_set_with_predictions)[names(test_set_with_predictions) == "lwr"] <- "Lower_Bound"
names(test_set_with_predictions)[names(test_set_with_predictions) == "upr"] <- "Upper_Bound"

# Scatter plot of Actual vs. Predicted with Confidence Intervals
library(ggplot2)
ggplot(test_set_with_predictions, aes(x = number_of_billionaires, y = Predicted_Billionaires)) +
  geom_point(color = 'blue', size = 2) + # Points for actual vs predicted
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") + # 1:1 line
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), width = 0.2, color = 'red') +
  labs(
    title = "Actual vs Predicted Number of Billionaires with Confidence Intervals",
    x = "Actual Number of Billionaires",
    y = "Predicted Number of Billionaires"
  ) +
  theme_minimal()

```

```{r}
# Use the model to predict number of billionaires and the Prediction Interval for the new data
predicted_billionaires <- predict(model_multiple, test_set, 
                                  interval = "prediction", level = 0.95)
predicted_billionaires
```

```{r}
# Combine actual and predicted values into the test set
test_set_with_predictions <- cbind(test_set, predicted_billionaires)

# Rename prediction columns for clarity
names(test_set_with_predictions)[names(test_set_with_predictions) == "fit"] <- "Predicted_Billionaires"
names(test_set_with_predictions)[names(test_set_with_predictions) == "lwr"] <- "Lower_Bound"
names(test_set_with_predictions)[names(test_set_with_predictions) == "upr"] <- "Upper_Bound"

# Scatter plot of Actual vs. Predicted with Prediction Intervals
library(ggplot2)
ggplot(test_set_with_predictions, aes(x = number_of_billionaires, y = Predicted_Billionaires)) +
  geom_point(color = 'blue', size = 2) + # Points for actual vs predicted
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") + # 1:1 line
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), width = 0.2, color = 'red') +
  labs(
    title = "Actual vs Predicted Number of Billionaires with Prediction Intervals",
    x = "Actual Number of Billionaires",
    y = "Predicted Number of Billionaires"
  ) +
  theme_minimal()

```

### Predicting number of billionaires in Malaysia for 2024

```{r}
new_data <- data.frame(gdp_country = 9e10, industry_diversity = 10)
predictions <- predict(model_multiple, new_data, interval = "prediction")
print(predictions)
```

```{r}
# Set the actual number of billionaires for Malaysia
actual_billionaires <- 20

# Create a comparison table
comparison <- data.frame(
  Country = "Malaysia",
  Actual = actual_billionaires,                   # Actual number of billionaires (from Forbes.com)
  Predicted = predictions[1],                    # Predicted number of billionaires
  Lower_Confidence = predictions[2],             # Lower bound of confidence interval
  Upper_Confidence = predictions[3]            # Upper bound of confidence interval
)

# Rename columns for clarity
colnames(comparison) <- c("Country", "Actual (forbes.com)", "Predicted", "Lower Confidence", "Upper Confidence")

# Print the comparison table
print(comparison)

```

### Visualisation

```{r}
library(ggplot2)

# Generate data for the model's prediction range
prediction_data <- data.frame(
  gdp_country = seq(min(new_billionaires_clean$gdp_country), max(new_billionaires_clean$gdp_country), length.out = 100),  # Adjust range
  industry_diversity = 10  # Keep constant
)

# Predict with intervals
prediction_data <- cbind(
  prediction_data,
  predict(model_multiple, newdata = prediction_data, interval = "prediction")
)

# Plot
ggplot(prediction_data, aes(x = gdp_country, y = fit)) +
  geom_line(color = "blue", size = 1.2) +  # Predicted trend
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, fill = "blue") +  # PI
  geom_point(data = new_data, aes(x = gdp_country, y = predictions[1, "fit"]), color = "red", size = 3) +
  labs(
    title = "Prediction of Number of Billionaires in Malaysia (2024)",
    x = "GDP (Country)",
    y = "Predicted Number of Billionaires",
    caption = "Blue line: Predicted trend | Shaded: Prediction Interval | Red: Malaysia 2024"
  ) +
  theme_minimal()

```
